---
title: "Computational Social Science"
subtitle: "A Graduate & Undergraduate Course"
author: "Course Instructor"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(kableExtra)
```

---

## Welcome

This course sits at the intersection of **sociology, political science, economics, psychology, communication, public health, criminology, and geography** — united by computational methods and a commitment to understanding the social world.

It is built on three commitments:

> **Disciplinary pluralism** — no single discipline owns social science
> **Methodological pluralism** — quantitative rigor and qualitative depth are complementary
> **Computational power with human judgment** — code is the tool; theory and ethics are the compass

---

## Course Flow

```
Epistemology & Foundations (01)
        ↓
Qualitative Roots & Mixed Methods (02)
        ↓
Measurement — EFA & CFA (03)
        ↓
Structure — SEM & Causal Inference (04)
        ↓
Typologies — LCA, LPA, RCA, CCA (05)
        ↓
Language — Text Analysis (06)
        ↓
Relationships — Network Analysis (07)
        ↓
Space — Spatial Social Science (08)
        ↓
Integration: Your Own Research (09)
```

---

## Modules

```{r module-table}
tibble::tribble(
  ~Module, ~Title, ~Methods, ~Packages, ~Link,
  "01", "Foundations, Epistemology & Working with Social Data",
        "CSS overview, ethics, data types", "`gssr`, `tidyverse`",
        "[Open](modules/module_01_foundations.html)",
  "02", "Qualitative Roots & the Mixed-Methods Bridge",
        "CGT, QCA, content analysis", "`tidytext`, `QCA`",
        "[Open](modules/module_02_qualitative.html)",
  "03", "Measurement — Exploratory and Confirmatory Factor Analysis",
        "EFA, CFA, measurement invariance", "`psych`, `lavaan`",
        "[Open](modules/module_03_efa_cfa.html)",
  "04", "SEM, Regression, and Causal Inference",
        "Path models, mediation, DAGs, DiD", "`lavaan`, `dagitty`, `did`",
        "[Open](modules/module_04_sem_causal.html)",
  "05", "Latent Class and Latent Profile Analysis",
        "LCA, LPA, RCA, GoM, Q-method", "`poLCA`, `tidyLPA`, `corclass`",
        "[Open](modules/module_05_latent.html)",
  "06", "Computational Text Analysis",
        "LDA, STM, embeddings, BERT", "`quanteda`, `stm`, `text2vec`",
        "[Open](modules/module_06_text.html)",
  "07", "Social Network Analysis",
        "Centrality, communities, ERGM", "`igraph`, `statnet`",
        "[Open](modules/module_07_networks.html)",
  "08", "Spatial Social Science",
        "GIS, LISA, spatial regression", "`sf`, `spdep`, `tidycensus`",
        "[Open](modules/module_08_spatial.html)",
  "09", "Integration, Multi-Method Pipelines, and Capstone",
        "Multi-method design, capstone", "Everything",
        "[Open](modules/module_09_integration.html)"
) |>
  kbl(col.names = c("Module", "Title", "Key Methods", "Primary Packages", ""),
      escape = FALSE) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE) |>
  column_spec(1, bold = TRUE, width = "4em") |>
  column_spec(4, monospace = TRUE) |>
  column_spec(5, width = "5em")
```

---

## Exercises

```{r exercise-table}
tibble::tribble(
  ~Module, ~Title, ~Link,
  "01", "Foundations of Computational Social Science",
        "[Open](exercises/ex_01_foundations.html)",
  "02", "Qualitative Roots & Mixed-Methods Bridge",
        "[Open](exercises/ex_02_qualitative.html)",
  "03", "Exploratory & Confirmatory Factor Analysis",
        "[Open](exercises/ex_03_efa_cfa.html)",
  "04", "Structural Equation Modeling & Causal Inference",
        "[Open](exercises/ex_04_sem_causal.html)",
  "05", "Latent Class, Profile, Relational & Correlational Class Analysis",
        "[Open](exercises/ex_05_latent.html)",
  "06", "Text Analysis & Computational Language Methods",
        "[Open](exercises/ex_06_text.html)",
  "07", "Network Analysis",
        "[Open](exercises/ex_07_networks.html)",
  "08", "Spatial Social Science",
        "[Open](exercises/ex_08_spatial.html)"
) |>
  kbl(col.names = c("Module", "Title", ""),
      escape = FALSE) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE) |>
  column_spec(1, bold = TRUE, width = "4em") |>
  column_spec(3, width = "5em")
```

---

## Key Datasets

| Dataset | Source | Modules |
|---------|--------|---------|
| General Social Survey (GSS) | NORC / `gssr` R package | 01, 03, 05 |
| ANES 2020 Time Series | electionstudies.org | 01, 04, 05 |
| IPIP Big Five (`bfi`) | `psych` package | 03 |
| World Values Survey Wave 7 | worldvaluessurvey.org | 03, 05 |
| UN General Debate Corpus | Harvard Dataverse | 06 |
| US Inaugural Addresses | `quanteda` package | 06 |
| Zachary's Karate Club | `igraphdata` package | 07 |
| Florentine Families | `ergm` package | 07 |
| ACS (Census) | tidycensus + API key | 08 |

---

## Getting Started

**Step 1:** Open `css_course.Rproj` in RStudio
**Step 2:** Run `source("R/install_packages.R")`
**Step 3:** Run `source("R/download_data.R")`
**Step 4:** Open a module and click **Knit**

---

*Built with R, RMarkdown, and a deep commitment to open science.*
*All materials: CC BY 4.0 | Code: MIT License*
